---
title: "Chapter 3 Exercise 2: Titanic"
author: "Mary Wojtas"
date: "November 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
## (0) Load the data in RStudio
```{r}
library(readr)
titanic_original <- read_csv("titanic_original.csv")
newtitanic <- titanic_original
```
##(1) Port of embarkation: Find missing values and replace them with S
```{r}
which(is.na(newtitanic$embarked))
print(newtitanic$embarked[169])
newtitanic <- newtitanic %>% 
  mutate(embarked = ifelse(is.na(embarked),"S",embarked))
print(newtitanic$embarked[169])
```
Mary's old code:   

for(i in 1:length(newtitanic\$embarked)){  
  if(is.na(newtitanic\$embarked[[i]]) == TRUE){  
    newtitanic\$embarked[[i]] <-c("S")}}  

##(2) Age: Calculate the mean of the Age column and use that value to populate the missing values
### 1st: Calculating the average Age
```{r}
avgage <- mean(newtitanic$age[1:1310], na.rm = TRUE)
print(avgage)
```
### 2nd:  Use that value to populate the missing values in the Age column
```{r}
newtitanic <- newtitanic %>%
  mutate(age = ifelse(is.na(age),avgage,age))
```
Other ways to populate the missing values in the age column?  
1. You can drop the rows with the missing values  
2. You can just populate with NA if they aren't already populated with NA  

Mary's old code:  
Convert avgage to a character because newtitanic\$age is a vector  

avgage <- as.characer(avgage)  

Next, replace the missing values with this new mean 

for(i in 1:length(newtitanic\$age)){  
  if(is.na(newtitanic\$age[[i]])== TRUE){  
    newtitanic\$age[[i]] <- c(avgage)}}  

##(3) Lifeboat: Fill empty values in the boat column with a dummy value: None
```{r}
newtitanic <- newtitanic %>%
  mutate(boat = ifelse(is.na(boat), "NONE", boat))
```

##(4) Cabin: 
Many passengers don't have a cabin number associated with them.  
Does it make sense to fill missing cabin numbers with a value?  
Maybe not.  This could be a category by which to classify survival.  
What does the missing value mean?  
It could mean any of the following:  
* The person was a stow-away  
* The person was part of the crew and not a passenger
* The person was a passenger but didn't have a cabin assignment

### Create a bew column "has_cabin_number" which has 1 if there is a cabin number, and 0 otherwise.
```{r}
newtitanic <- newtitanic %>%
  mutate(has_cabin_number = ifelse(is.na(cabin), 0, 1))
```

Mary's old code:  

for(i in 1:length(newtitanic\$has_cabin_number)){  
  if(!is.na(newtitanic\$has_cabin_number[[i]]) == TRUE){  
    newtitanic\$has_cabin_number[[i]] <- c("1")}}  

for(i in 1:length(newtitanic\$has_cabin_number)){  
  if(is.na(newtitanic\$has_cabin_number[[i]]) == TRUE){  
    newtitanic\$has_cabin_number[[i]] <- c("0")}}  

##(5) Output the cleaned up data sheet

titanic_clean <- newtitanic  
titanic_clean <- as.data.frame(titanic_clean)  
titanic_clean <- write_csv(titanic_clean,"~/Documents/RProjects/Ch3_Exer2_Titanic/titanic_clean.csv")  








